The Heavy Photon Search Test Detector

### 2013-10-28 hps-testrun-nim.tex
Updating SVT DAQ section and fixing figure centers.

\subsubsection{Support Structure}

@@

The total power consumption of the five APV25 chips mounted on 
the hybrid readout board is about 2~W. Since heating from the leakage current is only significant at a 
single small spot on the sensor, the 
dominant heat load on the sensor, even after irradiation, is radiant heat from the inside wall of the vacuum 
chamber, less than 0.5~W per sensor. The total power consumption budget of about 40~W is removed by a 
water/glycol mixture circulated through a flexible manifold attached to the cupper tubes in the cooling blocks. The total mass flow is about XXXXg/s to operate the sensors at XXXX degrees C.

@@

\subsection{SVT Digitization, Event Building and Data Transmission}
\label{sec:svtdaq}
The SVT Data Acquisition (DAQ) is a SLAC-standard system using consisting of a single ATCA crate with 
two Cluster On Board (COB) cards and a 10~Gb/s switch~\cite{Larsen:2011zb}.  
The COBs connect to Rear Transition 
Modules (RTM) housing 14-bit ADCs that accept input data from the APV25 chips.  
The COB house FPGA-based 
Data Processing Modules (DPM) and the Trigger Interface (TI) to the ECal trigger.  This system is capable 
of 20~kHz operation, whereas upgrades for the full experiment pushes this up to 50~kHz. 
Power is provided to each hybrid using CAEN power supplies from the decommissioned CDF SVXII 
detector.

{\bf  This needs to be expanded. I should talk about the data reduction and event handling. }

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. 
The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svtdaq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[h]
	\begin{center}
	\includegraphics[width=6cm]{figures/08062012_run1351_samples_vs_amplitude}
    	\caption{{\small The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}}
	\label{fig:pulse_shape}
	\end{center}
\end{figure}
These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}. 
The signal to noise ratio is XXXX.
\begin{figure}[h]
	\begin{center}
	\includegraphics[width=6cm]{figures/run1351_mip_small.pdf}
    	\caption{{\small The cluster charge distribution for hits associated with a track. }}
	\label{fig:cluster_pulse}
	\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit time 
with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter due to clock 
phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[h]
	\begin{center}
%	\includegraphics[width=7cm]{figures/top_track_time.pdf}
	\includegraphics[width=6cm]{figures/t0_resolution.pdf}
	\caption{\small{
%	Track time distribution (left) and cluster time residual (right). The track time is measured relative to the APV25 clock. The width of the distribution is due to trigger jitter (24~ns jitter in the tracker readout clock, plus 16~ns jitter in the trigger system). 
	The cluster time residual is for a representative sensor relative to the track time.}}
	\label{fig:tracktime}
	\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) the extracted $t_0$ resolution is 
2.6~ns. This is somewhat worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate with 
the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

While occupancy was slightly larger than expected, good agreement between data and simulation was 
found after taking into account dead or noisy channels. The hit efficiency, see Fig.~\ref{fig:hit_efficiency}, 
was measured to be above 98\% and fairly uniform across the SVT.
\begin{figure}[h]
	\begin{center}
    	\includegraphics[width=7cmh]{figures/hit_efficiency_vs_layer.pdf}
        \caption{{\small The hit reconstruction efficiency as a function of detector layer.}} 
	\label{fig:hit_efficiency}
	\end{center}
\end{figure}

The spatial resolution of similar microstrip sensors is well established by test beam data, against which the 
charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise (S/N) and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ration above 20, as demonstrated here, 
is relatively constant at approximately 6~$\mu$m for tracks that are close to normal to the sensors as in HPS.

@@

\subsubsection{Avalanche Photodiode Readout and Preamplifiers}
\label{sec:apd}
Avalanche Photodiodes (APD) S8664-55 produced by Hamamatsu Corp. are used as photodetectors. 
They have a $5\times5$~mm$^2$ active area and quantum efficiency of about 75\%. Single APDs were 
centered on the back ends of the crystals using MeltMount 1.7 thermal plastic adhesive. 
The low? amplitude signals of APDs are amplified using custom?made preamplifier boards. 
Figure~\ref{fig:ecal-module} shows a schematic view of the module assembly.
\begin{center}
{\small
\begin{figure}[]
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
\end{figure}
}
\end{center}
The gain uniformity of the APDs is on the order of XXXX\%; with APD high voltage bias varying by as 
much as 100V. Since only a limited number of high voltage channels were available during the test, 10 
channels were grouped together with an attempt to group modules with similar gains.

@@

We calibrate gain of the individual ECal channels using the SVT measurement of track momentum and 
comparison to Monte Carlo simulation.  We disable all SVT and ECal channels in the simulation that were 
inoperable or noisy in the test run, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well; then we use a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[ht]
	\includegraphics[width=7cm]{figures/ecalgainplots_corr_sim}
	\includegraphics[width=7cm]{figures/gains}
	\caption{\small{Weighted E/p from Monte Carlo simulation (top), calibrated values of gain in units of MeV 
	per ADC count (bottom).}}
	\label{fig:gains}
\end{figure}

@@

\subsubsection{Flash ADC Readout Board}
\label{sec:fadc}
The analog signals from the individual APD's of the ECal (shaped and amplified as described in 
Sec.~\ref{sec:aid}) are input to a single channel on the 16-channel JLab FADC250 VXS module (FADC), 
shown in Fig. ~\ref{fig:fadc}. 
\begin{center}
{\small
\begin{figure}[t]
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
\end{figure}
}
\end{center}
Three 20-slot VXS crates are needed to accommodate the system: one for each half of the ECal with 221 
channels and one for the Muon System with a total of 232 channels.

@@

For the readout path, every FADC has the following parameters:
 \begin{itemize}
 \item the number of samples integrated before the signal crossed threshold ($NSB$), 
 \item the number of samples integrated after the signal crossed threshold ($NSA$),
 \item the readout threshold, measured in ADC counts.
 \end{itemize}
The number of samples for a given channel integration is the sum of $NSB+NSA$ samples. It is a fixed gate width pulse integration with no pedestal subtraction where the sum is stored in a 17-bit register for readout (pedestal subtraction happens offline). 
\begin{center}
{\small
\begin{figure}[]
\includegraphics[width=7cm]{figures/hps_trigger_data}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
\end{figure}
}
\end{center}
For the trigger path, every channel has, in addition to $NSB$ and $NSA$: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item trigger threshold, measured in ADC counts, 
 \item a pedestal, 
 \item a conversion factor (gain) that converts the ADC counts to energy in MeV (with 13 bits:  from 0 to 8191~MeV), 
 \item an energy discriminator threshold (minimum energy cutoff).
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The pedestal value is subtracted from the integrated sum over $NSB+NSA$ samples and 
converted to MeV units using a supplied gain conversion factor. The energy discriminator can be used to cut 
off low energy pulses before reporting to the Crate Trigger Processor. 
The values reported to the Crate Trigger Processor are the 13-bit pulse energy and the time at which the 
pulse crossed the threshold. Data for every channel is sent every 32~ns (if there is no hit a 0 energy pulse is 
sent) which sets a worst case double pulse resolution of 32~ns for individual channels, but less if pulses 
occur in adjacent 32~ns windows.

@@

\subsection{Trigger}
\label{sec:trigger}
The trigger system is designed to efficiently select \ee{} events by 
using information from the ECal . The trigger looks for time coincidences of clusters in the top and bottom half 
of the ECal.  The trigger system can be broken down into three sections (see Fig.~\ref{fig:hps_trigger_cal}):
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. Sends the measured pulse energy and arrival time to the CTP.
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. The cluster energy, arrival time, and hit pattern are sent to the SSP.
 \item SSP (cluster pair finding): searches for time coincidence of pairs of clusters from the top and bottom half of the ECal and applies topological selections.
 \end{itemize}
\begin{center}
{\small
 \begin{figure}[t]
 \includegraphics[width=7cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes the detector channel signals and sends them for cluster finding in the CTP. The CTP clusters are sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout the sub-detectors.}
 \label{fig:hps_trigger_cal}
 \end{figure}
}
\end{center}
The time coincidence window of pairs of clusters in the top and bottom half of the ECal are programmable 
with 4~ns resolution.

@@

To eliminate trigger bias in checking the trigger decision, we use a tag and probe method. To check trigger 
performance in one half of the ECal, we tag events where there was a trigger in the other half, and exactly one 
probe cluster in the ECal half under test. 
We then measure trigger efficiency (proportion of tagged events where there was a trigger) as a function of 
ADC counts and energy of the probe cluster.
These turn-on curves are shown for the top half of the ECal in Fig.~\ref{fig:turnon}. 
\begin{center}
{\small
\begin{figure}[ht]
	\includegraphics[width=7cm]{figures/top_turnon_adc}
	\includegraphics[width=7cm]{figures/top_turnon_e}
	\caption{Trigger turn-on as a function of probe cluster ADC counts (top) and probe cluster energy in MeV 
	(bottom). Both plots are for the top half of the ECal; bottom is similar. Energy is not corrected for sampling 
	fraction.}
\label{fig:turnon}
\end{figure}
}
\end{center}
The trigger threshold is seen to be 1280 ADC counts as expected. The threshold is not perfectly sharp in this 
analysis because of uncertainties in the conversion from readout to trigger hits but based on comparisons 
with Monte Carlo simulation we believe the trigger worked exactly as specified. 
The trigger threshold in terms of cluster energy is very uneven for two reasons; gain variations between 
different ECal crystals lead to threshold variations and the nonlinearity of the time-over-threshold integral 
means that the effective threshold is higher for clusters that span multiple crystals.
Overall the trigger appears to have functioned exactly as intended. Changes planned for the next run 
(constant integration window and per-crystal gain calibration constants for the trigger) will solve both of the 
issues that led to threshold variations in the test run.

\subsection{Data Acquisition and Online Computing}
For the ECal, every VXS crate contains a Readout Controller (ROC)
that collects digitized information, processes it, and sends it on to the Event Builder (EB). The ROC is a single 
blade Intel-based CPU module running DAQ software under CentOS Linux OS. For the SVT ATCA system, 
the ROC application runs on an embedded processor situated on the ATCA main board. The EB assembles 
information from the SVT, ECal and Muon System ROCs into a single event which is passed to the Event 
Recorder (ER) that writes it to a RAID5-based data storage system capable of handling up to 100 MB/s. The 
EB and other critical components run on multicore Intel-based multi-CPU servers. The DAQ network system is 
a Foundry router providing high-speed connections between the DAQ components and to the JLab 
computing facility. The SVT ROC, which must handle large data volumes, has a 10 Gbit/s link to the Foundry 
router, while a 1 Gbit/s link is adequate for the ECal and Muon System. A 10 Gbit/s uplink to the JLab 
computing facility is used for long-term storage.

>>>>

\subsubsection{Support Structure and Cooling}

@@

The total power consumption of the five APV25 chips mounted on 
the hybrid readout board is about 2~W. Since heating from the leakage current is only significant at a 
single small spot on the sensor, the 
dominant heat load on the sensor, even after irradiation, is radiant heat from the inside wall of the vacuum 
chamber, less than 0.5~W per sensor. The total power consumption budget of about 40~W is removed by a 
water/glycol mixture circulated through a flexible manifold attached to the cupper tubes in the cooling blocks. 
During the Test run the sensors where operated at around $23^{\circ}$C with a total mass flow of 
about XXXXg/s.

@@

\subsection{SVT Digitization, Event Building and Data Transmission}
\label{sec:svtdaq}
The SVT data acquisition (DAQ) based on the Reconfigurable Cluster Element (RCE) and Cluster 
Interconnect (CI) concept developed at SLAC as generic building blocks for DAQ systems in high energy 
physics {\bf [need reference]}. 
The RCE is a generic computational building block, housed on a separate daughter card called 
Data Processing Module (DPM), that are realized on an ATCA front board called the Cluster On Board (COB), 
see Fig.~\ref{fig:cob}.
 \begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/svt_daq_module}
	\caption{Picture of the SVT DAQ COB board with four data processing daughter cards visible.}
	\label{fig:cob}
}
\end{center}
\end{figure}
The Gen-I RCE used in the Test run consisted of a Virtex~5 FPGA with 1~GB of DDR3 RAM. 
The outputs of 12 SVT half-modules 
are digitized on the Rear-Transition-Module (RTM), a custom board on the back side of the ATCA crate, 
interfacing the HPS readout to the generic DAQ components on the COB. 
A pre-amplifier converts the APV25 differential current output to a different voltage output scaled to the sensitive 
range of a 14-bit ADC operating at the system clock of 41.667~MHz. 
The RTM is organized into four sections with each section supporting three SVT half-module hybrids (15 
APV25 channels). The RTM also includes a 4-channel Fiber Optic module and supporting logic which can be 
used to interface to the JLab trigger system supervisor.
Each section of the RTM, up to three hybrids, is input to a DPM which apply thresholds for data reduction and 
organizes the sample data into UDP datagrams. The DPM also hosts an I2C controller used to configure and 
monitor the APV25 chips on those three hybrids. A single ATCA crate with two COB cards was used, one 
supporting four DPMs and one supporting 3 DPMs and one DPM that is configured as the trigger and data 
transmission module. The two COB cards and their DPMs are interconnected with a 10Gb/s switch 
card~\cite{Larsen:2011zb} which 
also hosts two 1Gb/s Ethernet interfaces to the external SVT DAQ PC.

The external PC supports three network interfaces; two standard 1Gb/s Ethernet and one custom low latency 
data reception card. The first Ethernet interface is used for slow control and monitoring of the 8 DPM modules. 
The second Ethernet interface serves as the interface to the JLAB data acquisition system. The third custom low 
latency network interface is used to receive data from the SVT ATCA crate and supports a low latency, reliable 
TTL trigger acknowledge interface to the trigger DPM. This PC hosts the SVT control and monitoring software 
as well as the JLab read out controller (ROC) application described in more detail in Sec.~\ref{sec:daq}.

This system is capable of operating up to 20~kHz and XXXX~MB/s which is well in line with the rates 
expected and observed during the HPS Test run.

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. 
The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svtdaq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/08062012_run1351_samples_vs_amplitude}
    	\caption{The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}

These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}. 
The signal to noise ratio is XXXX.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_mip_small.pdf}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit time 
with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter due to clock 
phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[h]
\begin{center}
{\small
%	\includegraphics[width=7cm]{figures/top_track_time.pdf}
	\includegraphics[width=6cm]{figures/t0_resolution.pdf}
	\caption{
%	Track time distribution (left) and cluster time residual (right). The track time is measured relative to the APV25 clock. The width of the distribution is due to trigger jitter (24~ns jitter in the tracker readout clock, plus 16~ns jitter in the trigger system). 
	The cluster time residual is for a representative sensor relative to the track time.}
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) the extracted $t_0$ resolution is 
2.6~ns. This is somewhat worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate with 
the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

While occupancy was slightly larger than expected, good agreement between data and simulation was 
found after taking into account dead or noisy channels. The hit efficiency, see Fig.~\ref{fig:hit_efficiency}, 
was measured to be above 98\% and fairly uniform across the SVT.
\begin{figure}[h]
\begin{center}
{\small
    	\includegraphics[width=7cmh]{figures/hit_efficiency_vs_layer.pdf}
        \caption{ The hit reconstruction efficiency as a function of detector layer.}
	\label{fig:hit_efficiency}
}
\end{center}
\end{figure}
The spatial resolution of similar microstrip sensors is well established by test beam data, against which the 
charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise (S/N) and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ration above 20, as demonstrated here, 
is relatively constant at approximately 6~$\mu$m for tracks that are close to normal to the sensors as in HPS.

@@

\subsubsection{Avalanche Photodiode Readout and Preamplifiers}
\label{sec:apd}
Avalanche Photodiodes (APD) S8664-55 produced by Hamamatsu Corp. are used as photodetectors. 
They have a $5\times5$~mm$^2$ active area and quantum efficiency of about 75\%. Single APDs were 
centered on the back ends of the crystals using MeltMount 1.7 thermal plastic adhesive. 
The low? amplitude signals of APDs are amplified using custom?made preamplifier boards. 
Figure~\ref{fig:ecal-module} shows a schematic view of the module assembly.
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}
The gain uniformity of the APDs is on the order of XXXX\%; with APD high voltage bias varying by as 
much as 100V. Since only a limited number of high voltage channels were available during the test, 10 
channels were grouped together with an attempt to group modules with similar gains.

@@

We calibrate gain of the individual ECal channels using the SVT measurement of track momentum and 
comparison to Monte Carlo simulation.  We disable all SVT and ECal channels in the simulation that were 
inoperable or noisy in the test run, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well; then we use a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/ecalgainplots_corr_sim}
	\includegraphics[width=7cm]{figures/gains}
	\caption{Weighted E/p from Monte Carlo simulation (top), calibrated values of gain in units of MeV 
	per ADC count (bottom).}
	\label{fig:gains}
}
\end{center}
\end{figure}

@@

\subsubsection{Flash ADC Readout Board}
\label{sec:fadc}
The analog signals from the individual APD's of the ECal (shaped and amplified as described in 
Sec.~\ref{sec:aid}) are input to a single channel on the 16-channel JLab FADC250 VXS module (FADC), 
shown in Fig. ~\ref{fig:fadc}. 
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Three 20-slot VXS crates are needed to accommodate the system: one for each half of the ECal with 221 
channels and one for the Muon System with a total of 232 channels.

@@

For the readout path, every FADC has the following parameters:
 \begin{itemize}
 \item the number of samples integrated before the signal crossed threshold ($NSB$), 
 \item the number of samples integrated after the signal crossed threshold ($NSA$),
 \item the readout threshold, measured in ADC counts.
 \end{itemize}
The number of samples for a given channel integration is the sum of $NSB+NSA$ samples. It is a fixed gate width pulse integration with no pedestal subtraction where the sum is stored in a 17-bit register for readout (pedestal subtraction happens offline). 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/hps_trigger_data}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}
For the trigger path, every channel has, in addition to $NSB$ and $NSA$: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item trigger threshold, measured in ADC counts, 
 \item a pedestal, 
 \item a conversion factor (gain) that converts the ADC counts to energy in MeV (with 13 bits:  from 0 to 8191~MeV), 
 \item an energy discriminator threshold (minimum energy cutoff).
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The pedestal value is subtracted from the integrated sum over $NSB+NSA$ samples and 
converted to MeV units using a supplied gain conversion factor. The energy discriminator can be used to cut 
off low energy pulses before reporting to the Crate Trigger Processor. 
The values reported to the Crate Trigger Processor are the 13-bit pulse energy and the time at which the 
pulse crossed the threshold. Data for every channel is sent every 32~ns (if there is no hit a 0 energy pulse is 
sent) which sets a worst case double pulse resolution of 32~ns for individual channels, but less if pulses 
occur in adjacent 32~ns windows.

@@

\subsection{Trigger}
\label{sec:trigger}
The trigger system is designed to efficiently select \ee{} events by 
using information from the ECal . The trigger looks for time coincidences of clusters in the top and bottom half 
of the ECal.  The trigger system can be broken down into three sections (see Fig.~\ref{fig:hps_trigger_cal}):
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. Sends the measured pulse energy and arrival time to the CTP.
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. The cluster energy, arrival time, and hit pattern are sent to the SSP.
 \item SSP (cluster pair finding): searches for time coincidence of pairs of clusters from the top and bottom half of the ECal and applies topological selections.
 \end{itemize}
 \begin{figure}[t]
\begin{center}
{\small
 \includegraphics[width=7cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes the detector channel signals and sends them for cluster finding in the CTP. The CTP clusters are sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
The time coincidence window of pairs of clusters in the top and bottom half of the ECal are programmable 
with 4~ns resolution.

@@

To eliminate trigger bias in checking the trigger decision, we use a tag and probe method. To check trigger 
performance in one half of the ECal, we tag events where there was a trigger in the other half, and exactly one 
probe cluster in the ECal half under test. 
We then measure trigger efficiency (proportion of tagged events where there was a trigger) as a function of 
ADC counts and energy of the probe cluster.
These turn-on curves are shown for the top half of the ECal in Fig.~\ref{fig:turnon}. 
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turnon_adc}
	\includegraphics[width=7cm]{figures/top_turnon_e}
	\caption{Trigger turn-on as a function of probe cluster ADC counts (top) and probe cluster energy in MeV 
	(bottom). Both plots are for the top half of the ECal; bottom is similar. Energy is not corrected for sampling 
	fraction.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger threshold is seen to be 1280 ADC counts as expected. The threshold is not perfectly sharp in this 
analysis because of uncertainties in the conversion from readout to trigger hits but based on comparisons 
with Monte Carlo simulation we believe the trigger worked exactly as specified. 
The trigger threshold in terms of cluster energy is very uneven for two reasons; gain variations between 
different ECal crystals lead to threshold variations and the nonlinearity of the time-over-threshold integral 
means that the effective threshold is higher for clusters that span multiple crystals.
Overall the trigger appears to have functioned exactly as intended. Changes planned for the next run 
(constant integration window and per-crystal gain calibration constants for the trigger) will solve both of the 
issues that led to threshold variations in the test run.

\subsection{Data Acquisition and Online Computing}
\label{sec:daq}
For the ECal, every VXS crate contains a Readout Controller (ROC)
that collects digitized information, processes it, and sends it on to the Event Builder (EB). The ROC is a single 
blade Intel-based CPU module running DAQ software under CentOS Linux OS. For the SVT ATCA system, 
the ROC application runs on an embedded processor situated on the ATCA main board. The EB assembles 
information from the SVT, ECal and Muon System ROCs into a single event which is passed to the Event 
Recorder (ER) that writes it to a RAID5-based data storage system capable of handling up to 100 MB/s. The 
EB and other critical components run on multicore Intel-based multi-CPU servers. The DAQ network system is 
a Foundry router providing high-speed connections between the DAQ components and to the JLab 
computing facility. The SVT ROC, which must handle large data volumes, has a 10 Gbit/s link to the Foundry 
router, while a 1 Gbit/s link is adequate for the ECal and Muon System. A 10 Gbit/s uplink to the JLab 
computing facility is used for long-term storage.
